---
title: "ADP Tables"
date: "2024-08-28"
output: 
  officedown::rdocx_document:
    reference_docx: "reference-word-doc.docx"
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(data.table)
library(magrittr)
library(flextable)
library(officer)

knitr::opts_chunk$set(echo = FALSE)

# Setting for flextable outputs
set_flextable_defaults(font.family = "Times New Roman", font.size = 12) 

#' Load the tables, generated by `selection_rates.R`
load("tables.rdata")

#' [NOTE: Draft vs Final]
#' For the draft, portrait orientation is probably fine, but you can use the set the output pages to landscape 
#' orientation using the calls to `block_section()` that may be commented out for the draft.
```

\newpage

Table B-2.  Budget allocation and vessel participation. The preliminary budget through the observer fee and NMFS funds for monitoring (both observer and EM) is $4.4 million. The number of vessels participating is estimated as the number of unique vessels that fished within each stratum within 365 days prior to the completion of the analyses. Some vessels may fish in multiple strata (e.g., a fixed-gear vessel may fish within the at-sea observer fixed-gear GOA and BSAI strata.)

```{r table_b2}
table_b2.flex
#block_section(prop_section(page_size = page_size(orient = "portrait"), type = "continuous"))
```

\newpage

Table B-3.  Partial coverage allocation indices, including the estimated numbers of trips in a stratum ($N_{h}$), monitoring rates ($r_{h}$), number of observed or monitored trips ($n_{h}$), the proximity indices ($\widehat{T}_{h}$), variance scaling factor ($F_{h}$), and proximity allocation index ($\widehat{D}_{h}$). 

```{r table_b3}
table_b3.flex
#block_section(prop_section(page_size = page_size(orient = "landscape"), type = "continuous"))
```

\newpage

Table B-4.  Summary of monitoring rates. Estimated numbers of trips in a stratum ($N_{h}$), number of observed or monitored trips ($n_{h}$), observed or monitored days ($d_{h}$), and monitoring rates ($r_{h}$) resulting from the deployment sampling design described in the text for 2025.

```{r table_b4}
table_b4.flex
#block_section(prop_section(page_size = page_size(orient = "portrait"), type = "continuous"))
```