---
title: "Between-Trip Variance Evaluation Metric"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r start, include=FALSE}
# Get packages
library(data.table)
library(flextable)
library(tidyverse)

# Get data
load("source_data/2024_Draft_ADP_data.rdata")
rm(list=setdiff(ls(), c("trips_melt", "efrt", "work.data")))

# Settings
knitr::opts_chunk$set(echo = FALSE)
options(scipen = 9999)
```

```{r wrangle}
# Convert optimization metrics to wide format
metrics <- dcast(trips_melt[, .(TRIP_ID, STRATA = strata_ID, METRIC = Metric, VALUE = Value)], TRIP_ID + STRATA ~ METRIC, value.var = "VALUE")

# Join optimization metric values with the most recent year of trips
dat <- metrics[efrt[ADP == 2022], on = .(TRIP_ID, STRATA), nomatch=0]

# Total number of trips per stratum
dat[, N := uniqueN(TRIP_ID), keyby = .(STRATA)]

# Get observed trip status
dat <- unique(work.data[, .(TRIP_ID, OBSERVED_FLAG)])[dat, on = .(TRIP_ID)]

# Sum metrics by columns of interest
dat <- dat[, .(chnk_psc = sum(chnk_psc), hlbt_psc = sum(hlbt_psc), discard = sum(discard)), by = .(TRIP_ID, STRATA, OBSERVED_FLAG, N)]

# Calculate between-trip standard deviation
dat <- dat[, .(sd_chnk = sd(chnk_psc[OBSERVED_FLAG == "Y"]), sd_hlbt = sd(hlbt_psc[OBSERVED_FLAG == "Y"]), sd_discard = sd(discard[OBSERVED_FLAG == "Y"])), by = .(STRATA, N)]

# Join with the full range of sample sizes for each stratum  
dat <- dat[, .(n = 2:N), by = .(STRATA)][dat, on = .(STRATA)]

# Convert to long format  
dat <- melt(dat, id.vars = c("STRATA", "n", "N"), variable.name = "metric", value.name = "sd")

# Calculate stratum-level standard error  
dat[, se := N * sqrt((N-n)/N) * sqrt(1/n) * sd]

# Recode strata and metric names
dat[, ":=" (STRATA = recode(STRATA, "EM_HAL" = "EM HAL", "EM_POT" = "EM POT", "EM_TRW_EFP" = "EM TRW EFP"), metric = recode(metric, "sd_chnk" = "Chinook PSC", "sd_hlbt" = "Halibut PSC", "sd_discard" = "Groundfish discards"))]

# Calculate monitoring rate
dat[, monitoring_rate := n/N]
```

In constructing the between-trip variance evaluation metric, we start by generating estimates of standard error at the stratum-level for each of the current optimization metrics using the formula $N\times\sqrt{\frac{N - n}{N} \frac{1}{n} \frac{\sum_{i=1}^{n} (x-\bar{x})^2}{n-1}}$:  
```{r plot}
ggplot(dat, aes(x = n, y = se)) +
  geom_line() +
  facet_grid(metric ~ STRATA, scales = "free") +
  theme_bw()
```

The data used for this calculation include zeros for trips that did not encounter any chinook PSC, halibut PSC, or groundfish discards. An assumption built into this approach is that all three metrics are measured without error at the trip-level, even though that is only true of chinook PSC. If we wanted to see what standard errors would result from two of the allocation schemes proposed in the Draft 2022 ADP (NMFS 2021), we could look at the selection rates that were proposed:  
```{r example_rates, message = FALSE}
adp_rates <- data.frame(STRATA = c("EM HAL", "EM POT", "EM TRW EFP", "HAL", "POT", "TRW"),
                        `Equal Rates` = c(30.00, 30.00, 33.33, 19.44, 19.44, 19.44),
                        `15% + Opt 0.95` = c(30.00, 30.00, 33.33, 18.21, 17.48, 28.10),
                         check.names = FALSE)

reshape2::melt(adp_rates, id.vars = "STRATA", variable.name = "Allocation scheme") %>% 
pivot_wider(names_from = STRATA, values_from = value) %>% 
flextable() %>% 
autofit()
```

And the standard errors associated with those selection rates:  
```{r example_errors}
se_adp <- dat %>% 
          left_join(adp_rates, by = "STRATA") %>% 
          melt(id.vars = c("STRATA", "metric", "N", "n", "monitoring_rate", "se"), measure.vars = c("Equal Rates", "15% + Opt 0.95"), variable.name = "Allocation scheme", value.name = "adp_rate") %>% 
          filter(round(adp_rate/100 * N) == n)

se_adp %>% 
select(STRATA, `Allocation scheme`, Metric = metric, se) %>% 
pivot_wider(names_from = STRATA, values_from = se) %>% 
# mutate(Total = rowSums(select_if(., is.numeric), na.rm = TRUE)) %>% 
mutate_if(is.numeric, round) %>% 
arrange(Metric, `Allocation scheme`) %>% 
flextable() %>% 
width(j = c(1,2), width = 1.5)
```

This table shows that the Equal Rates allocation strategy has a higher margin of error in Chinook PSC by 346 salmon (8%), in halibut PSC by 42 metric tons (15%), and in groundfish discards by 58 metric tons (2%) when compared to 15% + Opt 0.95. Although the simplifying assumptions built into this evaluation metric mean that these are not the actual margins of error we would expect, the metric does show the relative impact on uncertainty at the trip-level that changes to selection rates have. If we wanted to condense the metric into one number for each design, we could scale relative to the Equal Rates design:   
```{r example_condense_1}
# se_adp %>% 
# select(`Allocation scheme`, Metric, Total) %>% 
# group_by(Metric) %>% 
# mutate(Score = round(Total / Total[`Allocation scheme` == "Equal Rates"], 2)) %>% 
# select(-Total) %>% 
# flextable() %>% 
# autofit()
```
